### 2019-03-01 Seneca.tex
Markus' pass. A bunch of minor tweaks.

To facilitate model search and selection using the serverless architecture, 
we have developed Seneca, 
a framework for tuning the hyperparameters of machine learning applications 
in AWS Lambda. 
The Seneca pipeline consists of packaging, deployment, 
function optimization, and hyperparameter tuning. Figure~\ref{fig:seneca} 
shows the architecture of Seneca.  In the the upper-right-front, we
show the three inputs that Seneca expects from its users: (\textit{A}) a hyperparameter
configuration file, (\textit{B}) a dataset URL, and (\textit{C}) the lambda function of machine learning application. The configuration file specifies the range of values for each hyperparameter that the application expects. Seneca applies Cartesian product to all options in this configuration to construct the search space. The dataset URL refers to a valid dataset stored in the AWS Simple Storage 
Service (S3)~\cite{ref:awss3}.
%\footnote{\url{https://aws.amazon.com/s3/}}.

Seneca automatically builds and deploys an AWS Lambda application from the machine learning application specified.
To do so, Seneca launches a docker container that mirrors the
AWS Lambda execution environment, checks and installs
the machine learning application and any libraries it requires,
and compresses the application and uploads it to S3 (a work-around
for the 10MB AWS Lambda function size restriction).
Seneca constructs an AWS Lambda function from a template
that, when executed, will download the dataset and split it into a training and testing set, and construct, test, and evaluate a model using the application
and a set of hyperparameter values passed in by Seneca as arguments.
Users can specify the train/test split ratio that should be used by Seneca; the
default is 80\%/20\% for classification tasks.
The function returns a testing score. 
Upon completion of this process, the container deploys the function
to AWS Lambda using the
AWS Command Line Interface~\cite{ref:awscli}
%\footnote{\url{https://aws.amazon.com/cli/}} 
and the developers credentials.

@@

The cost of using AWS Lambda (i.e. \texttt{compute charge}) 
is the billed duration (execution time rounded up to the
nearest 100ms)~\cite{ref:pricing}
%~\footnote{\url{https://aws.amazon.com/lambda/pricing/}}
multiplied by the allocated memory of each invoked function.
One goal of our work is to optimize memory use by these
applications in order to reduce cost, and to investigate
the trade offs of doing so.

Currently, allocated memory for a Lambda function can be set
from 128MB to 3008MB in increments of 64MB.
AWS documentation~\cite{ref:lambdalimits}
%~\footnote{\url{https://docs.aws.amazon.com/lambda/latest/dg/resource-model.html}} 
states that Lambda allocates CPU to functions proportionally to allocated memory size,
as is done for general purpose AWS EC2 instance types.

@@

The Seneca optimizer first configures and invokes the function using a user-defined payload.  From this run, Seneca obtains the \texttt{maximum memory used}
by function reported by AWS Cloudwatch,
which it uses as its starting point.
Seneca then defines two double-ended queues 
(\textit{deque}) of length N, to store 
\texttt{allocated memory} and \texttt{compute charge} data of different invocations. While the current 
allocated memory is less than or equal to 3008 MB, 
the optimizer reconfigures and invokes the function 
using the next increment for memory allocation.  It 
calculates the compute charge for each invocation 
using current allocated memory and billed duration.

We employ two exit conditions. The first is when the  compute charge monotonically increases across \textit{deque}. The second is when the increase in slope is greater than a threshold. When the optimizer finds that both conditions hold, it pops the left-most value from \textit{deque} and configures the function to use that value for allocated memory for all future invocations. If these two conditions can not be satisfied through search, the allocated memory will be configured as the memory size that results in minimal compute charge within the \textit{deque}. After extensive experimentation, we find that $N=5$ and a slope threshold of 1 work best, but these values are configurable. In addition, this optimization can be turned on or off via a command line argument to Seneca.

>>>>

To facilitate model search and selection using the serverless
architecture, we have developed Seneca, a framework for tuning the
hyperparameters of machine learning applications in AWS Lambda.  The
Seneca pipeline consists of packaging, deployment, function
optimization, and hyperparameter tuning. Figure~\ref{fig:seneca} shows
the architecture of Seneca.  In the the upper-right-front, we show the
three inputs that Seneca expects from its users: (\textit{A}) a
hyperparameter configuration file, (\textit{B}) a dataset URL, and
(\textit{C}) the lambda function of the machine learning
application. The configuration file specifies a set of values for
each hyperparameter that the application expects. Seneca creates the
Cartesian product of all options in this configuration as
the search space. The dataset URL refers to a valid dataset stored in
the AWS Simple Storage Service (S3)~\cite{ref:awss3}.
%\footnote{\url{https://aws.amazon.com/s3/}}.

Based on the specified machine learning application, Seneca
automatically builds and deploys an AWS Lambda application by
launching a Docker container that mirrors the AWS Lambda execution
environment, checks and installs the machine learning application and
any libraries it requires, compresses the application and uploads it
to S3 (a
work-around for the 10MB AWS Lambda function size restriction).
Seneca constructs an AWS Lambda function from a template that, when
executed, will download the dataset and split it into a training and
testing set, and construct, test, and evaluate a model using the
application and a set of hyperparameter values passed in by Seneca as
arguments.  Users can specify the train/test split ratio that should
be used by Seneca; the default is 80\%/20\%.  The function returns a
testing score.  Upon completion of this process, the container deploys
the function to AWS Lambda using the AWS Command Line
Interface~\cite{ref:awscli}
%\footnote{\url{https://aws.amazon.com/cli/}} 
and the developers credentials.

@@

The cost of using AWS Lambda (i.e. \texttt{compute charge}) 
is the billed duration (execution time rounded up to the
nearest 100ms)~\cite{ref:pricing}
%~\footnote{\url{https://aws.amazon.com/lambda/pricing/}}
multiplied by the allocated memory of each invoked function.
One goal of our work is to optimize memory use of these
applications in order to reduce cost, and to investigate
the trade-offs of doing so.

Currently, allocated memory for a Lambda function can be set
from 128MB to 3008MB in increments of 64MB.
AWS documentation~\cite{ref:lambdalimits}
%~\footnote{\url{https://docs.aws.amazon.com/lambda/latest/dg/resource-model.html}} 
states that Lambda allocates CPU to functions  corresponding to allocated memory size,  
as is done for general purpose AWS EC2 instance types.

@@

The Seneca optimizer first configures and invokes the function using a user-defined payload.  From this run, Seneca obtains the \texttt{maximum memory used}
by the function as reported by AWS Cloudwatch,
and uses it as the starting point in its search.
Seneca then defines two double-ended queues 
(\textit{deque}) of length N, to store 
\texttt{allocated memory} and \texttt{compute charge} data of different invocations. While the current 
allocated memory is less than or equal to 3008 MB, 
the optimizer reconfigures and invokes the function 
using the next increment for memory allocation.  It 
calculates the compute charge for each invocation 
using current allocated memory and billed duration.

We employ two exit conditions. The first is when the  compute charge monotonically increases across \textit{deque}. The second is when the increase in slope is greater than a threshold. When the optimizer finds that both conditions hold, it pops the left-most value from \textit{deque} and configures the function to use that value for allocated memory for all future invocations. If these two conditions can not be satisfied during search, the allocated memory will be configured as the memory size that results in minimal compute charge within the \textit{deque}. After extensive experimentation, we find that $N=5$ and a slope threshold of 1 work best, but these values are configurable. In addition, this optimization can be turned on or off via a command line argument to Seneca.

### 2019-03-01 intro.tex
Markus' pass. A bunch of minor tweaks.

The scale and elasticity of cloud computing systems have 
fueled remarkable innovation and unprecedented commercial 
investment.  Cloud users ``rent'' virtualized resources (while sharing
the underlying physical resources) on a pay-per-use basis
in exchange for availability guarantees specified via service level
agreements (SLAs). Uniquely, cloud systems can be configured 
to add and remove (i.e. auto-scale) resources and services
automatically, based on the ``elastic'' resource requirements and service needs 
of executing applications.

To date however, clouds are used more for
enterprise services (object stores, databases, application servers, etc.)
than for elastic applications.
The reason is that it is challenging to configure complex
distributed systems for application use, 
and to leverage the auto-scaling that
clouds offer. To address this challenge, cloud providers 
have started to offer programming and execution environments
that preclude the need for server configuration,
under the \textit{serverless} moniker~\cite{ref:jonas2017occupy,ref:onesteptwostep,ref:peeking}.  
Serverless platforms automatically configure, manage, and scale applications
to significantly simplify cloud use.

>>>>

The scale and elasticity of cloud computing systems have 
fueled remarkable innovation and unprecedented commercial 
investment.  Cloud users ``rent'' virtualized resources (while sharing
the underlying physical resources) on a pay-per-use basis
in exchange for availability guarantees specified via service level
agreements (SLAs). Uniquely, cloud systems can be configured 
to add and remove (i.e. auto-scale) resources and services
automatically, based on the dynamic resource requirements and service needs 
of executing applications.

To date however, clouds are used more for
enterprise services (object stores, databases, application servers, etc.)
than for elastic applications.
The reason is that it is challenging to configure complex
distributed systems for application use, 
and to leverage the auto-scaling that
clouds offer. To address this challenge, cloud providers 
have started to offer programming and execution environments
that obviate the need for server configuration,
under the \textit{serverless} moniker~\cite{ref:jonas2017occupy,ref:onesteptwostep,ref:peeking}.  
Serverless platforms automatically configure, manage, and scale applications
to significantly simplify cloud use.

### 2019-03-01 related_work.tex
Markus' pass. A bunch of minor tweaks.

The serverless framework~\cite{ref:serverless_framework} provides
automated packaging and deployment for serverless functions across clouds. 
GammaRay~\cite{gammaray18} does so for AWS Lambda to insert profiling instrumentation.
The serverless framework uses 
CloudFormation~\cite{ref:cloudformation} for deployment in AWS Lambda, which introduces additional cost. 
The cloud infrastructure provisioning 
framework Terraform~\cite{ref:terraform}
also provides automated deployment of functions to serverless platforms.
%(however, without replicating the Lambda runtime, introducing deployment errors).
Seneca uses a local docker container to avoid  cost and overhead (vs these related
works), which guarantees execution compatibility for AWS Lambda. 
%In addition, AWS CodeDeploy~\cite{ref:codedeploy} serves as component in the continuous integration and delivery pipeline for serverless architectures, but it also introduces monetary cost overhead to do so.

Automated hyperparameter tuning is the focus of
many projects. Google Vizier~\cite{ref:vizier} provides a service for 
black-box optimization. Optunity~\cite{ref:claesen2014hyperparameter} 
and Hyperopt~\cite{ref:hyperopt}
provide a Python library for hyperparameter tuning. 
%This can be considered as precursory work of Hyperopt~\cite{ref:hyperopt}, 
%which is a more advanced Python library for serial and parallel hyperparameter 
%optimization over awkward search space. 
Hyperas~\cite{ref:hyperas} adds another 
abstraction layer to hyperopt to facilitate hyperparameter tuning for 
Keras~\cite{ref:keras}. However, no extant work of which 
we are aware leverages
serverless to perform hyperparameter tuning and memory optimization in parallel 
for ML applications.

>>>>

The serverless framework~\cite{ref:serverless_framework} provides
automated packaging and deployment for serverless functions across clouds. 
GammaRay~\cite{gammaray18} does so for AWS Lambda to insert profiling instrumentation.
The serverless framework uses 
CloudFormation~\cite{ref:cloudformation} for deployment in AWS Lambda, which introduces additional cost. 
The cloud infrastructure provisioning 
framework Terraform~\cite{ref:terraform}
also provides automated deployment of functions to serverless platforms.
%(however, without replicating the Lambda runtime, introducing deployment errors).
Seneca uses a local Docker container to avoid  cost and overhead (vs these related
works), which guarantees execution compatibility for AWS Lambda. 
%In addition, AWS CodeDeploy~\cite{ref:codedeploy} serves as component in the continuous integration and delivery pipeline for serverless architectures, but it also introduces monetary cost overhead to do so.

Automated hyperparameter tuning is the focus of
many projects. Google Vizier~\cite{ref:vizier} provides a service for 
black-box optimization. Optunity~\cite{ref:claesen2014hyperparameter} 
and Hyperopt~\cite{ref:hyperopt}
provide a Python library for hyperparameter tuning. 
%This can be considered as precursory work of Hyperopt~\cite{ref:hyperopt}, 
%which is a more advanced Python library for serial and parallel hyperparameter 
%optimization over awkward search space. 
Hyperas~\cite{ref:hyperas} adds another 
abstraction layer to hyperopt to facilitate hyperparameter tuning for 
Keras~\cite{ref:keras}. However, we are not aware of any work that
leverages serverless to perform hyperparameter tuning and memory
optimization in parallel for ML applications.

### 2019-03-01 results.tex
Markus' pass. A bunch of minor tweaks.

Prophet~\cite{ref:prophet} is an open source time series analysis library developed
by Facebook.  The input dataset we consider is a time series of view counts
of Peyton Manning's Wikipedia page (Dec. 2007--Jan. 2016).
The dataset exhibits both seasonality and a holiday effect (e.g. around 
super bowl games).  We use the first 6 years as the training set 
and the last 2 years as the testing set.  We use a cross-validation 
horizon (sliding window) of 1-year, and a period (sliding pace) 
of 180 days.  As such, Seneca performs three cross-validations for 2-year test range.

Prophet expects multiple hyperparameters.
\textit{growth} specifies linear or logistic trend model growth and \textit{prior scale} indicates the strength of the sparse prior probability. There are three prior scale hyperparameters for change point, holidays, and seasonality. Since Prophet uses a Fourier sum to estimate seasonality, the \textit{fourier order} is the number of terms in the partial Fourier sum. \textit{Seasonality mode} indicates that the effect of seasonality is either multiplicative or additive. Finally, the width of uncertainty intervals is set using \textit{interval width}.

@@

Hyperparameter configuration for Multi-Regression is a subset of input SBC time series.
Seneca considers all \texttt{$2^N - 1$} potential subsets (for $N$ input time series).
For this application, the default parameterization is the 
full set of input time series (8 in this case).
The test dataset is a time 
series of the outdoor temperature (ground truth) 
over the same period.  
The application makes 
predictions for each of these outdoor temperatures
using the regression coefficients constructed from the training set
for each new value in the test set.
%and computes and returns the MSE between predicted 
%and ground truth value pairs.

@@

The applications train a model on a random subset (80\%) of the data. Each then uses this model to predict the field from which each fruit originates for the remaining 20\%. To study the impact of random data split, we consider multiple 80\%/20\% splits in our evaluation.
\subsection{Empirical Methodology}

@@

Seneca deploys the applications automatically over AWS Lambda
and extracts execution time and memory use from 
AWS CloudWatch~\cite{ref:awscloudwatch} logs.
We compute monetary cost using the AWS Lambda pricing model~\cite{ref:pricing}.
Each function downloads the training/testing dataset 
of the application from AWS S3 upon function invocation. 
We do not consider the cost of dataset storage 
in our cost computations (it is very small).
%\footnote{\url{https://aws.amazon.com/cloudwatch/}}
%\footnote{\url{https://aws.amazon.com/lambda/pricing/}}
We also evaluate Seneca's automatic memory optimization capabilities.  To
do so we compare the execution performance and cost of the applications using
the maximum allocatable memory size to the performance and cost when run with
Seneca's automatically determined memory size. Even though \texttt{maximum memory used} reported by AWS CloudWatch can fluctuate, we 
have verified that the optimized allocated memory is sufficient for all 
hyperparameter configurations to complete successfully.  We have also verified
that the memory requirements across hyperparemeter settings do not vary 
significantly. We plan to consider applications for which hyperparameter settings require
different maximum memory sizes as part of future work.

@@

We first evaluate the quality of the output generated by each
ML applications when Seneca determines the hyperparameter settings. We
first show the results for the regression applications in
Table~\ref{tab:mse}.
The first row of data is the number of hyperparameter 
configurations that Seneca considers for each.
The last three rows show the MSE for the default, worst, and best performing (Seneca's recommendation)
hyperparameter configuation (lower is better).
Seneca reduces MSE by 22.56\%, 15.94\%, and 44.88\%, for Prophet, Multi-Regression, and XGBoost,
respectively, for the datasets and training methodologies that we consider.
Verses the worst case, Seneca reduces MSE by 82.62\%, 78.01\%, and 99.28\%, respectively.

@@

To investigate the potential impact of Seneca's 80/20 percent data split for the classification
applications, we next evaluate the quality of the output generated from 
each when we consider different 80/20 random splits.
To enable this, we run Seneca 30 times to obtain execution time, memory use, 
and best accuracy score. We report the
mean and standard deviation (in parentheses) 
for execution time and memory use across runs,
and the best accuracy score in
Table~\ref{tab:exec_memory}. Each pair of rows shows the results for two different random splits.
Our earlier results use input 1; this table adds results for a second, 80/20 
random split of the input (we also considered other random splits, which we omit for brevity, 
and the results are similar).  The performance and Seneca score is similar across splits. 
This result indicates that for these applications, 
users can repeatedly employ the recommended models for inference on other datasets or splits,
to amortize the cost of using Seneca.

>>>>

Prophet~\cite{ref:prophet} is an open source time series analysis library developed
by Facebook.  The input dataset we consider is a time series of view counts
of Peyton Manning's Wikipedia page (Dec. 2007--Jan. 2016).
The dataset exhibits both seasonality and a holiday effect (e.g. around 
super bowl games).  We use the first 6 years as the training set 
and the last 2 years as the testing set.  We use a cross-validation 
horizon (sliding window) of 1-year, and a period (sliding pace) 
of 180 days.  As such, Seneca performs three cross-validations for a 2-year test range.

Prophet expects multiple hyperparameters: \textit{growth} specifies linear or logistic trend model growth and \textit{prior scale} indicates the strength of the sparse prior probability. There are three prior scale hyperparameters for change point, holidays, and seasonality. Since Prophet uses a Fourier sum to estimate seasonality, the \textit{fourier order} is the number of terms in the partial Fourier sum. \textit{Seasonality mode} indicates that the effect of seasonality is either multiplicative or additive. Finally, the width of uncertainty intervals is set using \textit{interval width}.

@@

Hyperparameter configuration for Multi-Regression is a subset of input SBC time series.
Seneca considers all \texttt{$2^N - 1$} non-emtpy potential subsets (for $N$ input time series).
For this application, the default parameterization is the 
full set of input time series (8 in this case).
The test dataset is a time 
series of the outdoor temperature (ground truth) 
over the same period.  
The application makes 
predictions for each of these outdoor temperatures
using the regression coefficients constructed from the training set
for each new value in the test set.
%and computes and returns the MSE between predicted 
%and ground truth value pairs.

@@

The applications train a model on a random subset (80\%) of the data. Each then uses this model to predict the field from which each fruit originates for the remaining 20\%. To study the impact of random data split, we consider multiple (30 in our experiments) 80\%/20\% splits in our evaluation.
\subsection{Empirical Methodology}

@@

Seneca deploys the applications automatically over AWS Lambda
and extracts execution time and memory use from 
AWS CloudWatch~\cite{ref:awscloudwatch} logs.
We compute monetary cost using the AWS Lambda pricing model~\cite{ref:pricing}.
Each function downloads the training/testing dataset 
of the application from AWS S3 upon function invocation. 
We do not consider the cost of dataset storage 
in our cost computations (it is very small, under X cents).
%\footnote{\url{https://aws.amazon.com/cloudwatch/}}
%\footnote{\url{https://aws.amazon.com/lambda/pricing/}}
We also evaluate Seneca's automatic memory optimization capabilities.  To
do so we compare the execution performance and cost of the applications using
the maximum allocatable memory size to the performance and cost when run with
Seneca's automatically determined memory size. Even though \texttt{maximum memory used} reported by AWS CloudWatch can fluctuate, we 
have verified that the optimized allocated memory is sufficient for all 
hyperparameter configurations to complete successfully.  We have also verified
that the memory requirements across hyperparemeter settings do not vary 
significantly. We plan to consider applications for which hyperparameter settings require
different maximum memory sizes as part of future work.

@@

We first evaluate the quality of the output generated by each
ML applications when Seneca determines the hyperparameter settings. We
first show the results for the regression applications in
Table~\ref{tab:mse}.
The first row of data is the number of hyperparameter 
configurations that Seneca considers for each.
The last three rows show the MSE for the default, worst, and best performing (Seneca's recommendation)
hyperparameter configuation (lower is better).
Seneca reduces MSE by 22.56\%, 15.94\%, and 44.88\%, for Prophet, Multi-Regression, and XGBoost,
respectively, for the datasets and training methodologies that we consider.
Compared to the worst case, Seneca reduces MSE by 82.62\%, 78.01\%, and 99.28\%, respectively.

@@

To investigate the potential impact of Seneca's 80/20 percent data split for the classification
applications, we next evaluate the quality of the output generated from 
each when we consider different 80/20 random splits.
For that purpose, we run Seneca 30 times to obtain execution time, memory use, 
and best accuracy score. We report the
mean and standard deviation (in parentheses) 
for execution time and memory use across runs,
and the best accuracy score in
Table~\ref{tab:exec_memory}. Each pair of rows shows the results for two different random splits.
Our earlier results use input 1; this table adds results for a second, 80/20 
random split of the input (we also considered other random splits, which we omit for brevity, 
and the results are similar).  The performance and Seneca score is similar across splits. 
This result indicates that for these applications, 
users can repeatedly employ the recommended models for inference on other datasets or splits,
to amortize the cost of using Seneca.
